---
title: "Final Project: The Birthweight Social"
author: "Zhezheng Jin, Ngan Le, Kevin Liu, Duong Nguyen"
date: "Fall 2023"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
editor_options: 
  chunk_output_type: console
---

```{css, echo=FALSE}
/* Style for the main tabs */
.nav-tabs > li > a {
  background-color: #A3C4E5; 
  color: #5F5B57; 
}

/* Style for the active main tab */
.nav-tabs > li.active > a {
  background-color: #4276AA; 
  color: #ffffff; 
}

/* Style for the sub-tabs */
.tab-content .tab-pane .nav > li > a {
  background-color: #D2D2D2; 
  color: #3E3D3D; 
}

/* Style for the active sub-tab */
.tab-content .tab-pane .nav > li.active > a {
  background-color: #6D6D6D; 
  color: #ffffff; 
}
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(readxl)
library(rvest)
library(plotly)
library(knitr)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
library(survival)
library(kableExtra)
library(purrr)
library(tableone)
library(skimr)
```

# Motivation

_Provide an overview of the project goals and motivation_

- The motivation for a project exploring the correlation between low birth weight, premature birth rate, cancer mortality rate, cardiovascular disease hospitalization rate, asthma hospitalization rate and PM2.5 concentration, as well as socioeconomic, racial, education, and gender factors across the 62 counties of New York State, stems from a compelling public health imperative. 

- All the outcomes of interest are critical indicators of population health. By examining the impact of fine particulate pollution, this project aims to uncover environmental health disparities that disproportionately affect under-served communities. 

- It seeks to inform public health policy by highlighting the need for targeted interventions that could mitigate the adverse effects of air pollution on vulnerable populations, including pregnant women and newborns. The analysis of socioeconomic and racial factors will offer insights into the complex interplay between environment and social determinants of health, potentially guiding urban planning and healthcare resource allocation. 

- Furthermore, by considering gender-specific vulnerabilities,the project aims to provide a comprehensive overview of the risk landscape, fostering community awareness and prompting action to improve air quality and health equity in one of the most densely populated urban areas in the world.

# Related Work: KEVIN

_Anything that inspired you, such as a paper, a web site, or something we discussed in class._

# Initial Questions: KEVIN

_What questions are you trying to answer? How did these questions evolve over the course of the project? What new questions did you consider in the course of your analysis?_

# Data

_Source, scraping method, cleaning, etc._

## Data Import

Click `show` on the right to view the code chunk for data importing.
```{r message = FALSE}
lowbirthweight <- read_csv("csv_NYC_lowbirthweight.csv")
pm2_5 <- read_csv("pm2.5.csv")
edu_NY <- read_excel("Edu_NY.xlsx")
race_NY <- read_excel("Race_NY.xlsx")
HHincome_NY <- read_excel("HHincome_NY.xlsx")
Age_NY <- read_excel("Age_NY.xlsx")
Sex_NY <- read_excel("Sex_NY.xlsx")
health <- read_excel("chir_current_data.xlsx")
uscounties <- read_csv("uscounties.csv") #Simplemaps.com
```

## Data Wrangling {.tabset .tabset-fade}

### PM2.5 Dataset  
```{r warning=FALSE}
pm2_5 <- pm2_5 %>% 
  janitor::clean_names()%>%
  select (county,value) %>%
  rename (annual_pm2.5 = "value")
```
The dataset is obtained from EPH Tracking Website from CDC (https://ephtracking.cdc.gov/DataExplorer/). This has `r nrow(pm2_5)` rows and `r ncol(pm2_5)` columns of data. In which, 2 variables are:  
  - `county`: NY county name   
  - `annual_pm2.5`: annual estimated PM2.5 concentration at each county (ug/m^3)

### Demographic Dataset {.tabset .tabset-fade}
The following set of demographic datasets is obtained from Census Reporter webpage (https://censusreporter.org/).

#### Education Attainment
```{r, warning=FALSE}
edu <- edu_NY %>% 
  janitor::clean_names()%>%
  mutate (percentage_high_education = (bach_male+master_male+prof_male+doct_male+bach_female+master_female+prof_female+doct_female)/total) %>%
  filter(str_detect (name, " County, NY")) %>% 
  mutate(county = str_replace (name, " County, NY", "")) %>%
  select(county,percentage_high_education) %>%
  mutate(county = str_replace (county, "St.", "St")) %>%
  mutate(county = str_replace (county, "Stuben", "Steuben"))%>%
  select(county,percentage_high_education)
```  
This has `r nrow(edu)` rows and `r ncol(edu)` columns of data. In which, 2 variables are:  
  - `county`: NY county name   
  - `percentage_high_education`: percentage of population who finished a higher education level (higher than a bachelor degree)

#### Ethnicity
```{r warning=FALSE}  
race_non_hisp_white <- race_NY %>% 
  janitor::clean_names()%>%
  filter( x1 == "White Non-Hispanic") %>% 
  pivot_longer(
    albany_county_ny : yates_county_ny,
    names_to = "county",
    values_to = "percent_non_hisp_white") %>% 
  select (county,percent_non_hisp_white) %>%
  separate(county, into = c("county", "x"), sep = "_county_") %>%
  select (county,percent_non_hisp_white ) %>%
  mutate(county = str_replace (county, "_", " ")) %>%
  mutate_at(vars(county), str_to_title)

race_non_hisp_black <- race_NY %>% 
  janitor::clean_names()%>%
  filter( x1 == "Black Non-Hispanic") %>% 
  pivot_longer(
    albany_county_ny : yates_county_ny,
    names_to = "county",
    values_to = "percent_non_hisp_black") %>% 
  select (county,percent_non_hisp_black) %>%
  separate(county, into = c("county", "x"), sep = "_county_") %>%
  select (county,percent_non_hisp_black ) %>%
  mutate(county = str_replace (county, "_", " ")) %>%
  mutate_at(vars(county), str_to_title)

race_hisp_white <- race_NY %>% 
  janitor::clean_names()%>%
  filter( x1 == "White Hispanic") %>% 
  pivot_longer(
    albany_county_ny : yates_county_ny,
    names_to = "county",
    values_to = "percent_hisp_white") %>% 
  select (county,percent_hisp_white) %>%
  separate(county, into = c("county", "x"), sep = "_county_") %>%
  select (county,percent_hisp_white ) %>%
  mutate(county = str_replace (county, "_", " ")) %>%
  mutate_at(vars(county), str_to_title)

race_hisp_black <- race_NY %>% 
  janitor::clean_names()%>%
  filter( x1 == "Black Hispanic") %>% 
  pivot_longer(
    albany_county_ny : yates_county_ny,
    names_to = "county",
    values_to = "percent_hisp_black") %>% 
  select (county,percent_hisp_black) %>%
  separate(county, into = c("county", "x"), sep = "_county_") %>%
  select (county,percent_hisp_black ) %>%
  mutate(county = str_replace (county, "_", " ")) %>%
  mutate_at(vars(county), str_to_title)

race_merge <- race_non_hisp_white %>%
  inner_join(race_non_hisp_black, by = "county") %>%
  inner_join(race_hisp_white, by = "county") %>%
  inner_join(race_hisp_black, by = "county")

race_merge <- race_merge %>% 
  mutate (percent_other = 1 - percent_non_hisp_white - percent_non_hisp_black - percent_hisp_white - percent_hisp_black)
```
This has `r nrow(race_merge)` rows and `r ncol(race_merge)` columns of data. In which, 6 variables are:  
  - `county`: NY county name   
  - `percent_non_hisp_white`: percentage of population who are identified as Non-Hispanic White  
  - `percent_non_hisp_black`: percentage of population who are identified as Non-Hispanic Black    
  - `percent_hisp_white`: percentage of population who are identified as Hispanic White  
  - `percent_hisp_black`: percentage of population who are identified as Hispanic Black  
  - `percent_other`: percentage of population who are identified as any other ethinic groups    


#### Household Income
```{r warning=FALSE}
income <- HHincome_NY %>%
  janitor::clean_names() %>% 
  filter (x1 == "percent_high_income") %>%
   pivot_longer(
    albany_county_ny : yates_county_ny,
    names_to = "county",
    values_to = "percent_high_income") %>%
  select (county,percent_high_income) %>%
  separate(county, into = c("county", "x"), sep = "_county_") %>%
  select (county,percent_high_income ) %>%
  mutate(county = str_replace (county, "_", " ")) %>%
  mutate_at(vars(county), str_to_title)
```
This has `r nrow(income)` rows and `r ncol(income)` columns of data. In which, 2 variables are:  
  - `county`: NY county name   
  - `percent_high_income`: percentage of population who are at higher income households (>$75,000 annually)

#### Median Age
```{r warning=FALSE}
age <- Age_NY %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    albany_county_ny : yates_county_ny,
    names_to = "county",
    values_to = "median_age") %>%
  select (county,median_age) %>%
  separate(county, into = c("county", "x"), sep = "_county_") %>%
  select (county,median_age ) %>%
  mutate(county = str_replace (county, "_", " ")) %>%
  mutate_at(vars(county), str_to_title)
```
This has `r nrow(age)` rows and `r ncol(age)` columns of data. In which, 2 variables are:  
  - `county`: NY county name   
  - `median_age`: median age of the population at each county  

#### Sex
```{r warning=FALSE}
sex <- Sex_NY %>% 
  janitor::clean_names() %>% 
  filter (x1 == "Male:") %>%
  pivot_longer(
    albany_county_ny : yates_county_ny,
    names_to = "county",
    values_to = "percent_male") %>%
  select (county,percent_male) %>%
  separate(county, into = c("county", "x"), sep = "_county_") %>%
  select (county,percent_male ) %>%
  mutate(county = str_replace (county, "_", " ")) %>%
  mutate_at(vars(county), str_to_title)
```
This has `r nrow(sex)` rows and `r ncol(sex)` columns of data. In which, 2 variables are:  
  - `county`: NY county name   
  - `percent_male`: percentage of population who are identified as male  

#### Low Birthweight Data 
```{r warning=FALSE}
lowbirthweight <- lowbirthweight %>% 
  janitor::clean_names()%>%
  select (region_county, percentage)%>%
  rename (county = "region_county", percent_lowbirthweight = "percentage") 
```
This has `r nrow(lowbirthweight)` rows and `r ncol(lowbirthweight)` columns of data. In which, 2 variables are:  
  - `county`: NY county name   
  - `percent_lowbirthweight`: percentage of children being born being identified as low birth weight (<2,500g) (https://www.health.ny.gov/)

#### Health Indicator Dataset 
The following dataset is obtained from New York State Department of Health (https://www.health.ny.gov/) that contains 4 different health indicators that will complement with the low birthweight. These 5 will act as our outcomes of interest for our regression models.

```{r warning=FALSE}
health <- health %>%
  janitor::clean_names()%>%
  select (geographic_area,indicator_title,topic_area,rate_percent,measurement) %>%
  filter( str_detect (geographic_area, " County")) %>% 
  mutate (county = str_replace (geographic_area, " County", "")) %>%
  select (county, everything()) %>%
  select (-geographic_area) %>%
  filter(topic_area == "Cancer Indicators" | topic_area == "Respiratory Disease Indicators" | topic_area == "Cardiovascular Disease Indicators" | topic_area == "Maternal and Infant Health Indicators")
  
cancer <- health %>% 
  filter (topic_area == "Cancer Indicators") %>% 
  filter (indicator_title == "All cancer incidence rate per 100,000") %>% 
  select (-c(indicator_title, topic_area, measurement)) %>%
  mutate(rate_percent = as.numeric(rate_percent)/10) %>%
  rename (cancer_mortality_per_10k = "rate_percent")

resp <- health %>%
  filter (topic_area == "Respiratory Disease Indicators") %>%
  filter (indicator_title == "Asthma hospitalization rate per 10,000") %>% 
  select (-c(indicator_title, topic_area, measurement)) %>% 
  rename (asthma_hosp_rate_per_10k = "rate_percent")

cardio <- health %>%
  filter (topic_area == "Cardiovascular Disease Indicators") %>%
  filter (indicator_title == "Cardiovascular disease hospitalization rate per 10,000") %>% 
  select (-c(indicator_title, topic_area, measurement)) %>% 
  rename (cardio_hosp_rate_per_10k = "rate_percent")

maternal <- health %>%
  filter (topic_area == "Maternal and Infant Health Indicators") %>%
  filter (indicator_title == "Percentage of premature births with <37 weeks gestation") %>% 
  select (-c(indicator_title, topic_area, measurement)) %>% 
  rename (premature_percentage = "rate_percent")

```
They are:   
  - `cancer_mortality_per_10k`: percentage of cancer mortality per 100 thousands people in each NY county (Cancer Indicator)    
  - `asthma_hosp_rate_per_10k`: percentage of asthma hospitalization per 10 thousands people in each NY county (Respiratory Disease Indicator)     
  - `cardio_hosp_rate_per_10k`": percentage of cardiovascular-disease-related hospitalization per 10 thousands people in each NY county (Cardiovascular Disease Indicator)   
  - `premature_percentage`: percentage of children being born prematurely (<37  gestational weeks) in each NY county

### Merge dataset
Here we perform `inner_join()` to create 2 bigger datasets `health_merge` and `demographic_merge`. Then, we join them with our `lowbirthweight` & `pm2_5` tp make a finalized data frame called `merge`. And, we will use this for regression model. 
```{r warning=FALSE}
health_merge <- maternal %>% 
  inner_join(cancer, by = "county") %>%
  inner_join(resp, by = "county") %>% 
  inner_join(cardio, by = "county") 

demographic_merge <- age %>%
  inner_join(sex,  by = "county") %>%
  inner_join(income, by = "county") %>% 
  inner_join(race_merge, by = "county") %>% 
  inner_join(edu, by = "county") %>%
  mutate(county = str_replace (county, "St ", "St. "))
  

merge <- lowbirthweight %>% 
  inner_join(pm2_5, by = "county") %>%
  inner_join(health_merge, by = "county") %>% 
  inner_join(demographic_merge, by = "county") 

merge <- merge %>% 
  select (county, annual_pm2.5,everything())%>%
  mutate(asthma_hosp_rate_per_10k = as.numeric(asthma_hosp_rate_per_10k))%>%
  mutate(cardio_hosp_rate_per_10k = as.numeric(cardio_hosp_rate_per_10k)) %>%
  mutate(premature_percentage = as.numeric(premature_percentage))%>%
  mutate(cancer_mortality_per_10k = as.numeric(cancer_mortality_per_10k))

```

### Making Map File
```{r warning=FALSE}
uscounties <- uscounties %>% 
  filter (state_id == "NY") %>%
  select (county, lat, lng)

map <- merge %>% 
  inner_join(uscounties, by ="county")

write.csv(map, "NY_map.csv", row.names = FALSE)

```
This file is specifically made for the purpose of making map (in Shiny App).


# EDA {.tabset .tabset-fade}

_Visualizations, summaries, and exploratory statistical analyses. Justify the steps you took, and show any major changes to your ideas._

I CRASH BUT DANNY PLEASE SUMMARIZE() TOP 5 AND BOTTOM 5 COUNTIES FOR THE OUTCOMES GRAPH (AND LIST THEM OUT IF POSSIBLE THX <3 )

## Descriptive Summary
```{r, warning=FALSE}
merge_cont =
  merge %>% 
  select(-county)

skim_table <- skim(merge_cont)
numeric_summary <- skim_table %>%
  select(Variables = skim_variable, Nmiss = n_missing, Mean = numeric.mean, Median = numeric.p50, STD = numeric.sd)


kable(numeric_summary, "html", digits = 1) %>%
  kable_styling(full_width = FALSE, position = "center", font_size = 13) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(2, width = "3cm") %>%
  column_spec(3, width = "3cm") %>%
  column_spec(4, width = "3cm") %>%
  column_spec(5, width = "2cm")

#numeric_summary_df <- as.data.frame(numeric_summary)

#summary_plot =
#  plot_ly(numeric_summary_df, x = ~Mean, y = ~Variables, type = 'box', text = ~Variables, hoverinfo = 'mean', mean = "Mean") %>%
#  layout(title = 'Mean',
#         xaxis = list(title = 'Mean'),
#         yaxis = list(title = 'Variables'))

#summary_plot

```

## Percentage of High Income
Hover over the interactive graph or zoom in for more detailed information. 

```{r warning=FALSE}
merge %>%
 plot_ly(
    x = ~reorder(county, percent_high_income),
    y = ~percent_high_income,
    type = "bar",
    marker = list(color = "red1")
  ) %>%
  layout(
    title = "Percentage of High Income",
    xaxis = list(title = "County Name", categoryorder = "total descending"),
    yaxis = list(title = "Percentage"),
    barmode = "stack"
  )
```

The bar chart depicts the "Percentage of High Income" across 62 counties of New York State, ordered from the highest to the lowest percentage. Nassau county has the highest percentage of household with an annual income of more than $75,000 (categorized as high income), while that of the Bronx county is the lowest.

## Racial Composition 
Hover over the interactive graph or zoom in for more detailed information. 

```{r warning=FALSE}
race_plot <- merge %>% 
  select (county, percent_non_hisp_white, percent_non_hisp_black, percent_hisp_white, percent_hisp_black, percent_other) %>%
  pivot_longer(
    cols = starts_with("percent_"), names_to = "race", values_to = "percentage") 

race_plot%>%
  plot_ly(x = ~county, y = ~percentage, type = "bar",color = ~race,colors = "RdYlGn", hoverinfo = "y+name") %>% 
  layout(barmode = "stack",
         title = "Racial Composition in NY county",
         xaxis = list(title = "County"),
         yaxis = list(title = "Percentage (%)"))
```

This stacked bar chart displays the racial demographics across 62 New York counties. Each bar represents a county and is divided into segments that reflect the percentage of the county's population of the following racial categories: non-Hispanic White, non-Hispanic Black, Hispanic White, Hispanic Black, and Other. The majority of New York State population is non-Hispanic-White, which is also the major racial group in almost all counties, except in Bronx county and Queens county. 

## 5 Outcome Graphs {.tabset .tabset-fade}
The prevalence of our 5 outcomes of interest across 62 counties are illustrated in these 5 graphs below.

### Low birthweight Rate 
Hover over the interactive graph or zoom in for more detailed information. 

```{r warning=FALSE}
LBR_graph <- merge %>%
  drop_na() %>%
  plot_ly(
    x = ~county,
    y = ~percent_lowbirthweight,
    color = ~county,
    type = "scatter",
    mode = "markers"
  ) %>%
  layout(
    title = "Rate of Low Birthweight",
    xaxis = list(title = "County", tickangle = 90),
    yaxis = list(title = "Percentage of Low Birthweight"),
    showlegend = FALSE
  )

LBR_graph
```

### Premature Birth 
Hover over the interactive graph or zoom in for more detailed information. 

```{r warning=FALSE}
premature_graph <- merge %>%
  drop_na() %>%
  plot_ly(
    x = ~county,
    y = ~premature_percentage,
    color = ~county,
    type = "scatter",
    mode = "markers"
  ) %>%
  layout(
    title = "Rate of Premature Birth",
    xaxis = list(title = "County", tickangle = 90),
    yaxis = list(title = "Rate of Premature Birth"),
    showlegend = FALSE
  )

premature_graph
```


### Asthma Hospitalization 
Hover over the interactive graph or zoom in for more detailed information. 

```{r warning=FALSE}
asthma_graph <- merge %>%
  drop_na() %>%
  plot_ly(
    x = ~county,
    y = ~asthma_hosp_rate_per_10k,
    color = ~county,
    type = "scatter",
    mode = "markers"
  ) %>%
  layout(
    title = "Rate of Asthma Hospitalization",
    xaxis = list(title = "County", tickangle = 90),
    yaxis = list(title = "Rate per 10k"),
    showlegend = FALSE
  )

asthma_graph
```

### Cancer Mortality Rate 
Hover over the interactive graph or zoom in for more detailed information. 

```{r warning=FALSE}
cancer_graph <- merge %>%
  drop_na() %>%
  plot_ly(
    x = ~county,
    y = ~cancer_mortality_per_10k,
    color = ~county,
    type = "scatter",
    mode = "markers"
  ) %>%
  layout(
    title = "Rate of Cancer Mortality",
    xaxis = list(title = "County", tickangle = 90),
    yaxis = list(title = "Rate per 100k"),
    showlegend = FALSE
  )

cancer_graph
```

### Cardiovascular Rate 
Hover over the interactive graph or zoom in for more detailed information. 

```{r warning=FALSE}
cardio_graph <- merge %>%
  drop_na() %>%
  plot_ly(
    x = ~county,
    y = ~cardio_hosp_rate_per_10k,
    color = ~county,
    type = "scatter",
    mode = "markers"
  ) %>%
  layout(
    title = "Rate of Cardiovascular Disease Hospitalization",
    xaxis = list(title = "County", tickangle = 90),
    yaxis = list(title = "Rate per 10k"),
    showlegend = FALSE
  )

cardio_graph
```

### Outcomes Overlay
Hover over the interactive graph or zoom in for more detailed information. 

NOT SURE IF WE STILL NEED THIS!!!!!!!

```{r warning=FALSE}
fig <-  plot_ly()

fig <- fig %>% add_trace(x = ~merge$county, y = ~merge$percent_lowbirthweight, type = "scatter", mode = "markers", name = "Low Birthweight Rate")
fig <- fig %>% add_trace(x = ~merge$county, y = ~merge$premature_percentage, type = "scatter", mode = "markers", name = "Premature Birth Rate")
fig <- fig %>% add_trace(x = ~merge$county, y = ~merge$asthma_hosp_rate_per_10k, type = "scatter", mode = "markers", name = "Asthma Hospitalization Rate")
fig <- fig %>% add_trace(x = ~merge$county, y = ~merge$cancer_mortality_per_10k, type = "scatter", mode = "markers", name = "Cancer Mortality Rate")
fig <- fig %>% add_trace(x = ~merge$county, y = ~merge$cardio_hosp_rate_per_10k, type = "scatter", mode = "markers", name = "Cardiovascular Disease Hospitalization Rate")

fig <- fig %>% layout(
  title = "Overlay of 5 Outcomes",
  xaxis = list(title = "County", tickangle = 90),
  yaxis = list(title = "Rate"),
  showlegend = TRUE
)
fig
```


# Statistical Analysis 

_If you undertake formal statistical analyses, describe these in detail_

## Method
After cleaning the data, we then look at the effect that the annual concentration of pm 2.5 separately has on multiple outcomes among 62 counties in New York, including low birth weight rate, premature birth rate, cancer mortality rate, asthma-related hospitalization rate, and cardiovascular-disease-related hospitalization rate. 

Since all of the outcomes are measured as rates (i.e., continuous), we use __linear regression (LR) models__ to assess the effect of annual concentration of pm 2.5 on the outcomes of interests, adjusting for age, sex, ethnicity, education, income. In our models: 

- Age is reported as a continuous variable that reflects the median age of each county.

- Sex is reported as a continuous variable that reflects the percentage of male in each county. The variable that reflects the percentage of female in each county is not included in our model to avoid perfect multicollinearity.

- Ethnicity is reported as a continuous variable that reflects the percentage of different ethnicity groups in each county, including Hispanic-Black, Hispanic-White, non-Hispanic-Black, non-Hispanic-White, and Others (if Asian, Native American, or belong to two or more race). The variable that reflects the percentage of people who belong to the category Others is not included in our model to avoid perfect multicollinearity.

- Education is reported as a continuous variable that reflects the percentage of people who have obtained one or more higher education degrees in each county.

- Income is reported as a continuous variable that reflects the percentage of household of each county that have annual income exceeding $75,000.

For each of the five outcomes of interest, we first start with the full model:

_Outcome = Annual pm 2.5 concentration + median age + percentage of male + percentage of Hispanic-Black + percentage of Hispanic-White + percentage of non-Hispanic-Black + percentage of non-Hispanic-White + percentage of household income exceeding 75,000 USD + percentage of people obtained higher education_

We then perform __bidirectional stepwise selection__ on all five models and output the best models. The full model and the resulting best models after stepwise selection for each outcome are detailed below:

## LR models by Outcomes {.tabset .tabset-fade}

### Low Birth Weight Rate 

#### Stepwise regression 

__Full model__
```{r warning=FALSE, class.source = 'fold-show'}
lm_pm2.5_birthweight_adjusted <-lm(percent_lowbirthweight ~ annual_pm2.5 + median_age + percent_high_income + percent_non_hisp_white + percent_non_hisp_black + percent_hisp_white + percent_hisp_black + percentage_high_education + percent_male, data=merge)
```

__Stepwise selection__
```{r message=FALSE, class.source = 'fold-show'}
lm_pm2.5_birthweight_adjusted_best <- step(lm_pm2.5_birthweight_adjusted, direction = 'both', trace=FALSE)

summary(lm_pm2.5_birthweight_adjusted_best)%>%
  tab_model()
```


#### Interpretation

The best model that was outputted by the stepwise regression is 

_low birth weight = 11.53 - 0.11(median age) + 8.05 (percent_non_hisp_black)_

- On average, for a county whose median age is 0 and has 0% of non-Hispanic-Black, the expected low birth weight rate is 11.5% (no practical meaning). Since the p-value is smaller than 0.05, we have sufficient evidence to claim that the intercept is significantly different than 0.

- On average, for every unit increase of median age (1 year), the expected low birth weight rate of the county decreases by 0.11%. Since the p-value is smaller than 0.05, we have sufficient evidence to claim that the beta coefficient for median age is significantly different than 0.

- On average, for every unit increase in the percentage of non-Hispanic-Black, the expected low birth weight rate of the county increases by 8.05%. Since the p-value is smaller than 0.05, we have sufficient evidence to claim that the beta coefficient for percentage of non-Hispanic-Black is significantly different than 0.

The adjusted R-squared of 0.249 implies that 24.9% of the variation in the response variable can be explained by its linear relationship with the set of the 2 predictors (median age and percentage of non-Hispanic-Black).


### Premature Birth Weight Rate 

#### Stepwise regression 

__Full model__
```{r warning=FALSE, class.source = 'fold-show'}
lm_pm2.5_premature_adjusted <-lm(premature_percentage ~ annual_pm2.5 + median_age + percent_high_income + percent_non_hisp_white + percent_non_hisp_black + percent_hisp_white + percent_hisp_black + percentage_high_education + percent_male, data=merge)
```

__Stepwise selection__
```{r warning=FALSE, class.source = 'fold-show'}
lm_pm2.5_premature_adjusted_best <- step(lm_pm2.5_premature_adjusted, direction = 'both', trace=FALSE)

summary(lm_pm2.5_premature_adjusted_best)%>%
  tab_model()
```

#### Interpretation

The best model that was outputted by the stepwise regression is 

_premature birth weight = 4.88 + 0.08(median age) + 8.03 (percent_non_hisp_black)_

- On average, for a county whose median age is 0 and has 0% of non-Hispanic-Black, the expected low birth weight rate is 4.88% (no practical meaning). Since the p-value is smaller than 0.05, we have sufficient evidence to claim that the intercept is significantly different than 0.

- On average, for every unit increase of median age (1 year), the expected low birth weight rate of the county decreases by 0.11%. However, since the p-value is larger than 0.05, we do not have sufficient evidence to claim that the beta coefficient for median age is significantly different than 0.

- On average, for every unit increase in the percentage of non-Hispanic-Black, the expected low birth weight rate of the county increases by 8.03%. Since the p-value is smaller than 0.05, we have sufficient evidence to claim that the beta coefficient for percentage of non-Hispanic-Black is significantly different than 0.

The adjusted R-squared of 0.105 implies that 10.5% of the variation in the response variable can be explained by its linear relationship with the set of the 2 predictors (median age and percentage of non-Hispanic-Black).

### Cancer Mortality Rate {.tabset .tabset-fade}
Detailed below are the the initial model, the final model and the interpretation for each outcome.

#### Stepwise regression 

__Full model__
```{r warning=FALSE, class.source = 'fold-show'}
lm_pm2.5_cancer_adjusted <-lm(cancer_mortality_per_10k ~annual_pm2.5 + median_age + percent_high_income + percent_non_hisp_white + percent_non_hisp_black + percent_hisp_white + percent_hisp_black + percentage_high_education + percent_male, data=merge)
```

__Stepwise selection__
```{r warning=FALSE, class.source = 'fold-show'}
lm_pm2.5_cancer_adjusted_best<-step(lm_pm2.5_cancer_adjusted, direction = 'both', trace=FALSE)

summary (lm_pm2.5_cancer_adjusted_best)%>%
  tab_model()
```

#### Interpretation

The best model that was outputted by the stepwise regression is 

_cancer mortality rate = -1.85 + 13.27(annual pm 2.5) + 13.14(median age) - 213.49(percent high income) + 464.03(percent non hisp white) + 536.28(percent non hisp black) +  1023.82(percent hisp white) - 1787.43(percent hisp black) - 574.15(percent male)_

- On average, for a county whose annual pm 2.5 concentration is 0, median age is 0, has 0% of high income, has 0% of non-Hispanic-Black, has 0% of non-Hispanic-White, has 0% of Hispanic-Black, has 0% of Hispanic-White, has 0% of male, the expected low birth weight rate is -1.85 case per 100,000 people (no practical meaning). However, since the p-value is larger than 0.05, we do not have sufficient evidence to claim that the intercept is significantly different than 0.

- On average, for every unit increase of annual pm 2.5 concentration, the expected cancer mortality rate of the county increases by 13.27 deaths per 100,00 people. Since the p-value is smaller than 0.05, we have sufficient evidence to claim that the beta coefficient for annual pm 2.5 concentration is significantly different than 0. 

- On average, for every unit increase of median age (1 year), the expected cancer mortality rate of the county decreases by 13.14 cases per 100,000 people. Since the p-value is smaller than 0.05, we have sufficient evidence to claim that the beta coefficient for median age is significantly different than 0. 

- On average, for every unit increase in the percentage of high income, the expected cancer mortality rate of the county decreases by 213.49 cases per 100,000 people. Since the p-value is smaller than 0.05, we have sufficient evidence to claim that the beta coefficient for percentage of high income is significantly different than 0.

- On average, for every unit increase of percentage of non-Hispanic-White, the expected cancer mortality rate of the county increases by 464.03 deaths per 100,00 people. Since the p-value is smaller than 0.05, we have sufficient evidence to claim that the beta coefficient for percentage of non-Hispanic-White is significantly different than 0. 

- On average, for every unit increase of percentage of non-Hispanic-Black, the expected cancer mortality rate of the county increases by 536.28 deaths per 100,00 people. Since the p-value is smaller than 0.05, we have sufficient evidence to claim that the beta coefficient for percentage of non-Hispanic-Black is significantly different than 0. 

- On average, for every unit increase of percentage Hispanic-White, the expected cancer mortality rate of the county increases by 1023.08 deaths per 100,00 people. Since the p-value is smaller than 0.05, we have sufficient evidence to claim that the beta coefficient for percentage of Hispanic-White is significantly different than 0. 

- On average, for every unit increase of percentage Hispanic-Black, the expected cancer mortality rate of the county decreases by 1787.43 deaths per 100,00 people.  However, since the p-value is larger than 0.05, we do not have sufficient evidence to claim that the beta coefficient percentage of Hispanic-Black is significantly different than 0.

- On average, for every unit increase of percentage of male, the expected cancer mortality rate of the county decreases by 574.15 deaths per 100,00 people. However, since the p-value is larger than 0.05, we do not have sufficient evidence to claim that the beta coefficient for the percentage of male is significantly different than 0.

The adjusted R-squared of 0.780 implies that 70.8% of the variation in the response variable can be explained by its linear relationship with the set of the 8 predictors (annual pm 2.5 conc., median age, percentage of high income, percentage of 4 ethnicity groups, percentage of male).

### Asthma Hospitalization Rate {.tabset .tabset-fade}
Detailed below are the the initial model, the final model and the interpretation for each outcome.

#### Stepwise regression

__Full model__
```{r warning=FALSE, class.source = 'fold-show'}
lm_pm2.5_asthma_adjusted <-lm(asthma_hosp_rate_per_10k ~ annual_pm2.5 + median_age + percent_high_income + percent_non_hisp_white + percent_non_hisp_black + percent_hisp_white + percent_hisp_black + percentage_high_education + percent_male, data=merge)
```

__Stepwise selection__
```{r warning=FALSE, class.source = 'fold-show'}
lm_pm2.5_asthma_adjusted_best <- step(lm_pm2.5_asthma_adjusted, direction = 'both', trace=FALSE)

summary(lm_pm2.5_asthma_adjusted_best)%>%
  tab_model()
```

#### Interpretation

The best model that was outputted by the stepwise regression is 

_asthma hospitalization rate = 1.36 + 4.65(percent high income) + 13.52(percent non hisp black) + 366.67 (percent hisp black) - 5.31(percent high education)_

- On average, for a county who has zero percent of high income, has zero percent of non-Hispanic-Black, has zero percent of Hispanic-Black, has zero percent of high education, the expected asthma hospitalization rate is 1.85 case per 10,000 people (no practical meaning). Since the p-value is smaller han 0.05, we have sufficient evidence to claim that the intercept is significantly different than 0.

- On average, for every unit increase in the percentage of high income, the expected asthma hospitalization rate of the county increases by 4.65 cases per 10,000 people. Since the p-value is smaller than 0.05, we have sufficient evidence to claim that the beta coefficient for the percentage of high incomeis significantly different than 0.

- On average, for every unit increase of percentage of non-Hispanic-Black, the expected cancer mortality rate of the county increases by 13.52 deaths per 10,00 people. Since the p-value is smaller than 0.05, we have sufficient evidence to claim that the beta coefficient for the percentage of non-Hispanic-Black is significantly different than 0. 

- On average, for every unit increase of percentage of Hispanic-Black, the expected cancer mortality rate of the county increases by 366.67 deaths per 10,00 people. Since the p-value is smaller than 0.05, we have sufficient evidence to claim that the beta coefficient for the percentage of Hispanic-Black is significantly different than 0. 

- On average, for every unit increase of percentage of high education, the expected cancer mortality rate of the county decreases by 5.31 deaths per 10,00 people. Since the p-value is smaller than 0.05, we have sufficient evidence to claim that the beta coefficient for the percentage of higher education is significantly different than 0. 

The adjusted R-squared of 0.930 implies that 93.0% of the variation in the response variable can be explained by its linear relationship with the set of the 4 predictors (percentage of high income, percentage of non-Hispanic-Black, percentage of Hispanic-Black, percentage of higher education).

### Cardiovascular Disease Rate {.tabset .tabset-fade}
Detailed below are the the initial model, the final model and the interpretation for each outcome.

#### Stepwise regression

__Full model__
```{r warning=FALSE, class.source = 'fold-show'}
lm_pm2.5_cardio_adjusted <-lm(cardio_hosp_rate_per_10k ~annual_pm2.5 + median_age + percent_high_income + percent_non_hisp_white + percent_non_hisp_black + percent_hisp_white + percent_hisp_black + percentage_high_education + percent_male, data=merge)
```

__Stepwise selection__
```{r warning=FALSE, class.source = 'fold-show'}
lm_pm2.5_cardio_adjusted_best<-step(lm_pm2.5_cardio_adjusted, direction = 'both', trace=FALSE)

summary (lm_pm2.5_cardio_adjusted_best)%>%
  tab_model()
```

#### Interpretation

The best model that was outputted by the stepwise regression is 

_cardiovascular disease hospitalization rate = 157.52 + 6.97(annual pm 2.5) + 1.73(median age) + 117.26 (percent non hisp white) + 222.99(percent non hisp black) +  439.56(percent hisp white) - 148.61.43(percent high education) - 405.53(percent male)_

On average, for a county whose annual pm 2.5 concentration is 0, median age is 0, has zero percent of non-Hispanic-White, has zero percent of non-Hispanic-Black, has zero percent of Hispanic-White, has zero percent of high education, has zero percent of male, the expected cardiovascular disease hospitalization rate is 157.52 case per 10,000 people (no practical meaning). However, since the p-value is larger than 0.05, we do not have sufficient evidence to claim that the intercept is significantly different than 0.

- On average, for every unit increase of annual pm 2.5 concentration, the expected cardiovascular disease hospitalization rate of the county increases by 6.97cases per 10,000 people. However, since the p-value is smaller than 0.05, we do not have sufficient evidence to claim that the beta coefficient for annual pm 2.5 concentration is significantly different than 0. 

- On average, for every unit increase of median age (1 year), the expected cardiovascular disease hospitalization rate of the county decreases by 1.73 cases per 10,000 people. Since the p-value is smaller than 0.05, we have sufficient evidence to claim that the beta coefficient for median age is significantly different than 0. 

- On average, for every unit increase of percent of non-Hispanic-White, the expected cardiovascular disease hospitalization rate of the county increases by 177.26 cases per 10,000 people. However, since the p-value is larger than 0.05, we do not have sufficient evidence to claim that the beta coefficient for the percent of non-Hispanic-White is significantly different than 0. 

- On average, for every unit increase of percent of non-Hispanic-Black, the expected cardiovascular disease hospitalization rate of the county increases by 222.99 cases per 10,000 people. However, since the p-value is larger than 0.05, we do not have sufficient evidence to claim that the beta coefficient for percent of non-Hispanic-Black is significantly different than 0.

- On average, for every unit increase of percent of Hispanic-White, the expected cardiovascular disease hospitalization rate of the county increases by 439.56 cases per 10,000 people. However, since the p-value is larger than 0.05, we do not have sufficient evidence to claim that the beta coefficient for percent of Hispanic-White is significantly different than 0. 

- On average, for every unit increase of percent of higher education, the expected cardiovascular disease hospitalization rate of the county decreases by 148.61 cases per 10,000 people. Since the p-value is larger than 0.05, we have sufficient evidence to claim that the beta coefficient for percent of higher education is significantly different than 0. 

- On average, for every unit increase of percentage of male, the expected cardiovascular disease hospitalization rate of the county decreases by 405.53 cases per 10,000 people. However, since the p-value is larger than 0.05, we do not have sufficient evidence to claim that the beta coefficient for percentage male is significantly different than 0. 

The adjusted R-squared of 0.242 implies that 24.2% of the variation in the response variable can be explained by its linear relationship with the set of the 7 predictors (annual pm 2.5 concentration, median age, percent non-Hispanic-White, percent non-Hispanic-Black, percent Hispanic-White, percentage of higher education, percentage of male).

# Discussion: KEVIN

_What were your findings? Are they what you expect? What insights into the data can you make?_


